---
title: "wingen-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{wingen-vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, message = FALSE}
library(wingen)
library(terra)
library(raster)
library(viridis)
library(sf)
library(ggplot2)
```

## Background

wingen uses a moving window approach to create maps of genetic
diversity. The method and its rationale are described in [Bishop et al.
(2023)](http://doi.org/10.1111/2041-210X.14090) and on the [Methods
blog](https://methodsblog.com/2023/05/03/wingen-mapping-genetic-diversity-using-moving-windows/).

## Example

To demonstrate how wingen works, we will use a subset of the data from
the [Bishop et al. (2023)](http://doi.org/10.1111/2041-210X.14090)
simulation example. These simulations were created using Geonomics
([Terasaki Hart et al., 2022](https://doi.org/10.1093/molbev/msab175))
to generate a realistic landscape genomic dataset. In this simulation,
spatial variation in genetic diversity is produced by varying population
size and gene flow across the landscape via heterogeneous carrying
capacity and conductance surfaces. These surfaces are based on an
example digital elevation model of Tolkien's Middle Earth produced by
the Center for Geospatial Analysis at William & Mary ([Robert,
2020](https://scholarworks.wm.edu/asoer/3/)).

### Load Middle Earth example

The small Middle Earth example dataset used here contains four objects
which are loaded by `load_middle_earth_ex()`:

1.  `lotr_vcf` - a vcfR object containing the genetic data

2.  `lotr_coords` - a dataframe object containing sample coordinates

3.  `lotr_lyr` - a raster object of the landscape (higher values
    indicate greater connectivity/carrying capacity)

4.  `lotr_range` - a polygon outlining the "range" of the simulated
    species

```{r load example data, fig.width = 5, fig.height = 5}
load_middle_earth_ex()

# Genetic data
lotr_vcf

# Coordinates
head(lotr_coords)

# Raster data
lotr_lyr

# Map of data
plot(lotr_lyr, col = magma(100), axes = FALSE, box = FALSE)
points(lotr_coords, col = mako(1, begin = 0.8), pch = 3, cex = 0.5)
```

If users don't have a raster layer of their landscape, they can generate
one using their sample coordinates with the `coords_to_raster()`
function. The resolution of this raster can be tuned either with the
`agg` (to aggregate) or `disagg` (to disaggregate) arguments, or defined
using the `res` argument. The `res` argument can either be a single
value (e.g., 0.00833) or a vector of two values with the x and y
resolutions. The `buffer` argument can be used to add an edge to the
raster (i.e., buffer away from the coordinates).

```{r, fig.width = 5, fig.height = 5}
ex_raster1 <- coords_to_raster(lotr_coords, buffer = 1, plot = TRUE)
ex_raster2 <- coords_to_raster(lotr_coords, buffer = 1, agg = 2, plot = TRUE)
ex_raster3 <- coords_to_raster(lotr_coords, buffer = 1, disagg = 4, plot = TRUE)
ex_raster4 <- coords_to_raster(lotr_coords, buffer = 1, res = 10, plot = TRUE)
```

## Workflow

The workflow of wingen uses three main functions:

1.  `window_gd()`, `circle_gd()`, or `resist_gd()` to generate moving
    window maps of genetic diversity

2.  `krig_gd()` to use kriging to interpolate the moving window maps

3.  `mask_gd()` to mask areas of the maps from (1) and (2) (e.g., to
    exclude areas outside the study region)

## Different moving window calculations

There are three functions that can be used to generate moving window
maps:

1.  `window_gd()` uses a simple rectangular window with user-specified
    dimensions; this is the original method described in Bishop et al.
    2023

2.  `circle_gd()` uses a circular window with a user-specified radius

3.  `resist_gd()` uses a resistance distance-based window with a
    user-specified maximum distance (this can be thought of as a
    circular window with a radius that varies depending on the
    resistance of the landscape around it)

The main arguments that all three functions have in common are:

1.  `gen` - An object of type vcfR or a path to a vcf file with genotype
    data. The order of this file matters! The coordinate and genetic
    data should be in the same order, as there are currently no checks
    for this.

2.  `coords` - sf points, or a matrix or dataframe with two columns
    representing the coordinates of the samples. The first column should
    be x and the second should be y.

3.  `lyr` - A SpatRaster or RasterLayer which the window will move
    across to create the final map. In most cases, this will take the
    form of a raster of the study area.

4.  `stat` - The genetic diversity summary statistic to calculate.
    Current genetic diversity metrics that can be specified with `stat`
    include:

    -   `"pi"` for nucleotide diversity (default) calculated using
        `hierfstat::pi.dosage()` (only works for bi-allelic data).

    -   `"Ho"` for average observed heterozygosity across all sites

    -   `"allelic_richness"` for average number of alleles across all
        sites

    -   `"biallelic_richness"` for average allelic richness across all
        sites for a biallelic dataset (this option is faster than
        `"allelic_richness"`). When calculating `“biallelic_richness”`,
        users can choose to rarify allele counts (as in
        `hierfstat::allelic.richness()`) by setting
        `rarify_alleles = TRUE` (the default) or to use the raw allele
        counts by setting `rarify_alleles = FALSE`. We recommend
        performing allele count rarefaction (`rarify_alleles = TRUE`) if
        there are missing values in the genetic data, but for datasets
        with no missing data, it is faster to use the raw counts
        (`rarify_alleles = FALSE`).

    -   `"hwe"` for the proportion of sites that are not in
        Hardy--Weinberg equilibrium, calculated using `pegas`
        \link[pegas]{hw.test} at the 0.05 level (other alpha levels can
        be specified by adding the sig argument; e.g., `sig = 0.10`).

    -   `"basic_stats"` for a series of statistics produced by
        `hierfstat::basic.stats()` including mean observed
        heterozygosity (same as Ho), mean gene diversities within
        population (Hs), G ene diversities overall (Ht), and Fis
        following Nei (1987). Population-based statistics (e.g., FST)
        normally reported by `hierfstat::basic.stats()` are not included
        as they are not meaningful within the individual-based moving
        windows.

5.  `fact` - To decrease computational time, we provide the option to
    aggregate the input raster layer by some factor defined using the
    `fact` argument. Increasing `fact` will decrease the number of cells
    and thereby decrease the number of calculations, with the trade-off
    that the resolution of the output layers will decrease. Users should
    keep in mind that if they increase `fact`, they may simultaneously
    want to decrease `wdim` because the proportion of the landscape
    covered by the neighborhood matrix could otherwise increase
    substantially.

6.  `rarify` - Users have the option to perform rarefaction by setting
    the `rarify` argument to `TRUE.` If `rarify = TRUE`, users then
    define `rarify_n` as the number of samples to rarify to and
    `rarify_nit` as the number of iterations for rarefaction (e.g., if
    `rarify_n = 4` and `rarify_nit = 5`, for each sample set, four
    random samples will be drawn five times). Users can also set
    `rarify_nit = "all"`to use all possible combinations of samples of
    size `rarify_n` within the window (for example, if `rarify_n = 4`
    and the number of samples in the window is 5, all 20 possible
    combinations of samples will be used). As the window moves across
    the landscape, three things can occur based on the number of samples
    in the window: (1) if the number of samples is lower than
    `rarify_n`, genetic diversity is not calculated and a raster value
    of NA is assigned, (2) if the number of samples is equal to
    `rarify_n`, the genetic diversity statistic is calculated for those
    samples, (3) if the number of samples is greater than `rarify_n`,
    rarefaction is implemented and that set of samples is subsampled
    `rarify_nit` times to a size of `rarify_n` and the mean (or another
    summary statistic set using `fun`) of those `rarify_nit` iterations
    is used. Note that if the number of samples is lower than `min_n`, a
    raster value of NA will always be assigned, even if the number of
    samples is not less than `rarify_n` (i.e., if `min = 2` and
    `rarify_n = 1` and the sample size is 1 in the window, then an NA
    value will be assigned; to change this behavior so that genetic
    diversity is calculated, set `min_n = 1`)

    If `rarify = FALSE`, rarefaction is not performed and only steps (1)
    and (2) from above occur such that: (1) if the number of samples in
    the window is less than the `min_n` argument, genetic diversity is
    not calculated and a raster value of NA is assigned, and (2) if the
    number of samples is equal to or greater than `min_n`, the genetic
    diversity statistic is calculated for those samples. We highly
    encourage users to perform rarefaction as genetic diversity
    statistics are sensitive to sample size. The main benefit of not
    performing rarefaction is decreased computational time; however,
    this is not worth the trade-off in inaccuracy unless you are
    confident that there is no effect of rarefaction after performing
    your analysis with and without rarefaction.

------------------------------------------------------------------------

*Note:* Coordinates and rasters used in wingen should be in a projected
(planar) coordinate system such that raster cells are of equal sizes.
Therefore, spherical systems (including latitude-longitude coordinate
systems) should be projected prior to use. An example of how this can be
accomplished is shown below. If no CRS is provided, a warning will be
given and wingen will assume the data are provided in a projected
system.

```{r}
# First, we create example latitude-longitude coordinates and rasters

## Example raster:
lyr_longlat <- rast(
  ncols = 40, nrows = 40, xmin = -110, xmax = -90, ymin = 40, ymax = 60,
  crs = "+proj=longlat +datum=WGS84"
)

## Example coordinates:
coords_df <- data.frame(x = c(-110, -90), y = c(40, 60))
coords_longlat <- st_as_sf(coords_df, coords = c("x", "y"), crs = "+proj=longlat")

# Next, the coordinates and raster can be projected to an equal area projection, in this case the Goode Homolosine projection (https://proj.org/operations/projections/goode.html):
coords_eq <- st_transform(coords_longlat, crs = "+proj=goode")
lyr_eq <- project(lyr_longlat, "+proj=goode")

# Coordinates can be switched back to latitude-longitude the same way by replacing "goode" with "longlat"
```

### Run moving window calculations with `window_gd()`

To create a moving window map with a rectangular window, use
`window_gd()`. The main **additional** arguments to `window_gd()` are:

1.  `wdim` - Used to create the neighborhood matrix for the moving
    window based on the dimensions provided. This argument can either be
    set to one value (e.g., 3) which will create a square window (e.g.,
    3 x 3), or two values, which will create a rectangular window (e.g.,
    3 x 5). We encourage users to experiment with different values of
    `wdim` to determine the sensitivity of their results to this
    parameter. Ideally, `wdim` would be set with some knowledge of the
    study system in mind (e.g., the dispersal patterns and/or
    neighborhood size of the study organism). A preview of the window
    size relative to the landscape can be obtained using the
    `preview_gd()` function.

2.  `crop_edges` - Whether to crop out the cells along the edge of the
    raster that are not surrounded by a full window. Users may want to
    do so to avoid "edge effects" caused by incomplete windows along the
    borders of the raster. As wingen is relatively insensitive to sample
    size, edge effects are not likely to have a very strong effect on
    diversity estimates, so by default `crop_edges = FALSE`.

Before running `window_gd()`, users can preview the moving window and
the counts within each cell of the raster to get a sense of how big the
window is and what the density of counts looks like across their
landscape. Here, we provide the raster layer (`lotr_lyr`), the
coordinates (`lotr_coords`), the window dimensions (7), the aggregation
factor (3), and the minimum sample number (`min_n`). `min_n` will be
used to mask the sample count layer to show how much of the landscape
will be excluded due to low sample count. We specify `method = window`
for a preview of `window_gd()`, we will discuss other methods (i.e.,
`circle` and `resist` later).

```{r, fig.width = 6, fig.height = 5, cache = TRUE, warning = FALSE}
preview_gd(lotr_lyr,
  lotr_coords,
  method = "window",
  wdim = 7,
  fact = 3,
  sample_count = TRUE,
  min_n = 2
)
```

Next, we run the moving window function with our vcf, coordinates, and
raster layer. Here, we set the parameters to calculate pi, using a
window size of 7 x 7, an aggregation factor of 3, and rarefaction with a
rarefaction size of 2 (i.e., minimum sample size of 2), and 5
iterations.

The `L` argument is used in the calculation of pi. If `L = NULL`
(default), the function returns the sum over SNPs of nucleotide
diversity. Otherwise, the function returns the average nucleotide
diversity per nucleotide given the length L of the sequence. Users may
want to set `L` differently depending on amounts of missing data, for
example, so that missing sites are not considered in the calculation.

We then plot the genetic diversity layer (the first layer of the
produced raster stack) and the sample counts layer (the second layer).

*Note:* you will get a warning that no CRS is found for the provided
coordinates or raster and to check that the CRS for these objects match.
This is because these simulated data doesn't have a CRS, but we know
they match, so we can ignore this warning

```{r moving window, fig.width = 5, fig.height = 5, cache = TRUE, warning = TRUE, message = FALSE}
wgd <- window_gd(lotr_vcf,
  lotr_coords,
  lotr_lyr,
  stat = "pi",
  wdim = 7,
  fact = 3,
  rarify = TRUE,
  rarify_n = 2,
  rarify_nit = 5,
  L = 100
)
# The ggplot_gd function plots the genetic diversity layer
ggplot_gd(wgd, bkg = lotr_range) +
  ggtitle("Moving window pi") +
  # add coordinates
  geom_point(data = lotr_coords, aes(x = x, y = y), pch = 16)
```
```{r, fig.height = 5, fig.width = 5.5}
# The plot_count function plots the sample count layer
ggplot_count(wgd) +
  ggtitle("Moving window sample counts")
```

You can also create base R plots of your results using `plot_gd()` and
`plot_count()`

```{r, fig.width = 5, fig.height = 5}
plot_gd(wgd, bkg = lotr_range, main = "Moving window pi")

plot_count(wgd, main = "Moving window sample counts")
```

### Run moving window calculations with `circle_gd()`

To create a moving window map with a circular window, we can use
`circle_gd()`. The main **additional** arguments to `circle_gd()` are:

1.  `maxdist` - maximum geographic distance used to define the
    neighborhood; any samples farther than this distance will not be
    included (this can be thought of as the neighborhood radius). Can be
    provided as either (1) a single numeric value or (2) a SpatRaster
    where each pixel is the maximum distance to be used for that cell on
    the landscape (must be the same spatial scale as `lyr`). Similarly
    to `wdim`, we encourage users to experiment with different values of
    `maxdist` to determine the sensitivity of their results to this
    parameter and ideally set it with some knowledge of the study system
    in mind (e.g., the dispersal patterns and/or neighborhood size of
    the study organism).

2.  `distmat` - optional distance matrix output from `get_geodist()`; if
    not provided, `circle_gd()` will automatically use the
    `get_geodist()` function to calculate the distance matrix instead.

Like before, we can use the `preview_gd()` function to preview parameter
settings. This time we specify `method = "circle"` and need to provide
`maxdist`. Before doing so, we can calculate the distance matrix using
`get_geodist()` which we can then use in both `preview_gd()` and
`circle_gd()` to save time.

`get_geodist()` produces a distance matrix where the rows represent the
cells of the raster and the columns are the individuals. The values are
the geographic distance between each cell and each individual. Note that
the input layer for `get_geodist()` must have the same resolution as the
input layer for `preview_gd()` and `circle_gd()`. For example, if you
want to run `circle_gd()` with `fact = 3` and `lyr = lotr_lyr`, the
input layer to `get_geodist()` must be `aggregate(lotr_lyr, 3)`.
Alternatively, you can first create a new aggregated raster to be used
for all functions, as we have done below.

*Note:* if you want to get only the distances between your sample
coordinates, set `coords_only = TRUE`.

```{r, fig.width = 5, fig.height = 5, cache = TRUE, warning = FALSE}
# First, create a new aggregated raster
lotr_lyr5 <- aggregate(lotr_lyr, 5)

# Then calculate the distance matrix
distmat <- get_geodist(coords = lotr_coords, lyr = lotr_lyr5)

preview_gd(lotr_lyr5,
  lotr_coords,
  method = "circle",
  maxdist = 10,
  distmat = distmat,
  sample_count = TRUE,
  min_n = 2
)
```

Then, we can run `circle_gd()`:

```{r circle, fig.width = 5, fig.height = 5, cache = TRUE, warning = FALSE, message = FALSE}
cgd <- circle_gd(lotr_vcf,
  lotr_coords,
  lotr_lyr5,
  stat = "pi",
  maxdist = 10,
  distmat = distmat,
  rarify = FALSE,
  L = 100
)

ggplot_gd(cgd, bkg = lotr_range) +
  ggtitle("Circle moving window pi")
```

For `circle_gd()` (and `resist_gd()`), you can also define `maxdist`
with a raster to set the maximum distance (i.e., the radius) for each
cell. In this case, we can use the carrying capacity/conductance map
(`lotr_lyr5`) and multiply it by 100 so that the radius will range from
0 to 100 based on the value of the raster.

```{r variable circle, fig.width = 5, fig.height = 5, cache = TRUE, warning = FALSE, message = FALSE}
vcgd <- circle_gd(
  lotr_vcf,
  lotr_coords,
  lotr_lyr5,
  stat = "pi",
  maxdist = lotr_lyr5 * 100,
  distmat = distmat,
  rarify = FALSE,
  L = 100
)

ggplot_gd(vcgd, bkg = lotr_range) +
  ggtitle("Circle moving window pi")
```

### Run moving window calculations with `resist_gd()`

To create a moving window map with resistance distances, use
`resist_gd()`. The main **additional** arguments to `resist_gd()` are:

1.  `maxdist` - maximum cost distance used to define the neighborhood;
    any samples further than this cost distance will not be included
    (this can be thought of as the neighborhood radius, but in terms of
    cost distance). Can be provided as either (1) a single numeric value
    or (2) a SpatRaster where each pixel is the maximum distance to be
    used for that cell on the landscape (must be the same spatial scale
    as `lyr`). As stated before, we encourage users to experiment with
    different values of `maxdist` to determine the sensitivity of their
    results to this parameter.

2.  `lyr` - in addition to being the raster layer the window moves
    across, this will also serve as the conductivity layer used to
    calculate resistance distances (higher values should mean greater
    conductivity).

3.  `distmat` - optional distance matrix output from `get_resdist()`; if
    not provided, `resist_gd()` will automatically use the
    `get_resdist()` function to calculate the distance matrix instead.

`resist_gd()` takes a longer time to run and we recommend taking
advantage of parallelization, in particular for the calculation of the
distance matrix using `get_resdist()`. We also recommend first
calculating the distance matrix with `get_resdist()` and then providing
it to `resist_gd()` using the `distmat` argument as this allows for
`resist_gd()` to be run quickly multiple times (i.e., to test different
parameters) by avoiding repeating the costly distance calculations. Note
that `lyr` in both `get_resdist()` and in `resist_gd()` must have the
same spatial scale (e.g., the same resolution and extent).

As with `get_geodist()`, `get_resdist()` produces a distance matrix
where the rows represent the cells of the raster and the columns are the
individuals. The values are the resistance distance between each cell
and each individual. If you want to get only the distances between your
sample coordinates, set `coords_only = TRUE`.

Again, we can use the `preview_gd()` function to preview parameter
settings. Here, we specify `method = "resist"` and provide `maxdist`.

```{r, cache = TRUE, warning = FALSE, message = FALSE}
lotr_distmat <- get_resdist(coords = lotr_coords, lyr = lotr_lyr5)
```

```{r, fig.width = 6.5, fig.height = 5, cache = TRUE, warning = FALSE, message = FALSE}
preview_gd(lotr_lyr5,
  lotr_coords,
  method = "resist",
  maxdist = 60,
  distmat = lotr_distmat,
  sample_count = TRUE,
  min_n = 2
)
```

```{r resist, fig.width = 5, fig.height = 5, cache = TRUE, warning = FALSE, message = FALSE}
rgd <- resist_gd(lotr_vcf,
  lotr_coords,
  lotr_lyr5,
  distmat = lotr_distmat,
  stat = "pi",
  maxdist = 60,
  rarify = FALSE,
  L = 100
)

ggplot_gd(rgd, bkg = lotr_range) +
  ggtitle("Resistance moving window pi")
```

In addition to using continuous rasters of connectivity you can also use
rasters to designate passable and impassable areas. This can be useful
for designating boundaries you don't want windows to extend across
(e.g., mountain ranges) or extend between (e.g., rivers).

```{r, fig.width = 6.5, fig.height = 5, cache = TRUE, warning = FALSE}
# create layer
resist_lyr <- rast(aggregate(lotr_lyr, 5))

# create an impassable area down the middle coded as NA
resist_lyr[] <- 1
resist_lyr[ext(40, 60, -100, 0)] <- NA

# get resistance distances
distmat <- get_resdist(coords = lotr_coords, lyr = resist_lyr)

# plot preview
preview_gd(resist_lyr, lotr_coords, method = "resist", distmat = distmat, maxdist = 10)

# create moving window map
resist_gd <- resist_gd(
  lotr_vcf,
  lotr_coords,
  resist_lyr,
  maxdist = 10,
  distmat = distmat
)

ggplot_gd(resist_gd)
```

## Transforming the moving window maps

### Krige results

To produce smoother maps of genetic diversity, we provide the function
`krig_gd()` which creates a spatially interpolated raster from the
moving window raster produced by the moving window functions. This
function uses the `autoKrige()` function from the R package automap to
perform kriging on the moving window raster using an automatically
generated variogram. Note that the moving window raster stack, including
both the genetic diversity layer and the sample count layer, can be used
to generate kriged maps of both genetic diversity and sample count.

Kriging is performed by first transforming the moving window layer into
a set of coordinates with corresponding genetic diversity (or sample
count) values and then interpolating using these coordinates across the
grid provided. Because of this, it is important to keep in mind how the
coordinates from the moving window raster and the grid align. If the
resolution of the moving window raster is much lower than that of the
grid, there are fewer points for interpolation which can result in
grid-like artifacts during kriging.

To deal with this issue, users can either (1) resample their moving
window raster layers and grid layers to the same resolution using the
`resample` argument, or (2) manually disaggregate or aggregate either
the moving window or grid layers using the `agg_r`, `disagg_r`,
`agg_grd`, or `disagg_grd` arguments. Generally, if users want a
smoother resulting surface, a higher resolution grid layer should be
used (this can be accomplished by using the `disagg_grd` argument to
disaggregate the grid layer). The resampling, aggregation, and
disaggregation options currently only work if the object provided to
create the grid is a raster layer. Keep in mind that increasing the
resolution of the moving window layer (i.e., either by resampling or
disaggregating) can increase computational time substantially as this
increases the number of coordinates being used for kriging. This is also
the case for increasing the resolution of the grid layer, though to a
lesser extent.

To run this function, we provide the moving window raster stack output,
the indices of the layers we want to krige, and the raster layer to
interpolate across. We also disaggregate the original layer by a factor
of two to get a smoother output surface (users should play around with
this parameter). The output of this function is a raster stack of the
kriged input layers.

```{r krige results, fig.width = 5, fig.height = 5, cache = TRUE, warning = FALSE}
# Note: this step can take a little while
# index = 1 kriges the first layer in wgd (the genetic diversity layer)
kgd <- krig_gd(wgd, index = 1, grd = lotr_lyr, disagg_grd = 2)

ggplot_gd(kgd) +
  ggtitle("Kriged pi")
```

*Note:* By default, the kriging method is set to
`krig_method = "ordinary"` for ordinary/simple kriging (model formula:
\~ 1) which assumes that the mean and variance are constant across
space. Alternatively, `krig_method` can be set to `"universal"` for
universal kriging (model formula: \~ x + y), which allows the mean to
vary across locations, while variance is held constant.

For information on the differences between universal and ordinary
kriging, we recommend these articles:

-   [What are the different kriging
    models?](https://desktop.arcgis.com/en/arcmap/latest/extensions/geostatistical-analyst/what-are-the-different-kriging-models-.htm)

-   [Kriging
    interpolation](https://www.publichealth.columbia.edu/research/population-health-methods/kriging-interpolation)

Users can optionally get the full output from `autoKrige()` by setting `autoKrige_output = TRUE` in which case `krig_gd()` returns a list where the first element ("raster") is the kriged raster and second element ("autoKrige_output") is the autoKrige output. 

*Note:* `autoKrige()` does not work for non-projected systems (i.e.,
latitude-longitude) and will throw an error. See above example for how
latitude-longitude coordinates and rasters can be transformed.

*Note:* Kriging can produce values that fall outside the range of
possible values (e.g., genetic diversity values less than 0 or greater
than the possible maximum). By default, in order to ensure that the
values are bounded within the range of possible values, `krig_gd`
converts all values greater than the maximum of the input raster to that
maximum (`upper_bound = TRUE`), and all values less than the minimum of
the input raster to that minimum (`lower_bound = TRUE`). Users can turn
off this functionality by setting `lower_bound` and `upper_bound` to
`FALSE`. Users can also set their own custom `lower_bound` and
`upper_bound` values (e.g., for heterozygosity or pi, you may want to
set `lower_bound = 0` and `upper_bound = 1`).

### Mask results

Next, we mask the resulting kriged layers. Masking can be performed
using a variety of methods.

**Method 1:** Mask using the carrying capacity layer to exclude any
areas where the carrying capacity is lower than 0.1. Alternatively, one
could use a species distribution model or habitat suitability model to
exclude areas where the probability of presence is very low:

```{r mask results 1, fig.width = 5.5, fig.height = 5, warning = FALSE}
# Disaggregate lotr_lyr to make it the same resolution as kgd before masking
## Note: lotr_lyr is a RasterLayer which we convert to a SpatRaster with rast()
mask_lyr <- disagg(rast(lotr_lyr), 2)
mgd <- mask_gd(kgd, mask_lyr, minval = 0.1)

ggplot_gd(mgd) +
  ggtitle("Kriged & carrying capacity masked pi")
```

**Method 2:** Mask the layer using a species range map (in this case, an
sf polygon) to exclude areas falling outside the species range.

```{r mask results 2, fig.width = 5.5, fig.height = 5, warning = FALSE}
mgd <- mask_gd(kgd, lotr_range)

ggplot_gd(mgd) +
  ggtitle("Kriged & range masked pi")
```

**Method 3:** Mask the layer using a spatial Kernel Density Estimation
(KDE) of sample density to exclude areas with low sampling density using
the SpatialKDE package:

```{r, eval = FALSE}
# First, install and load the SpatialKDE package
if (!require("SpatialKDE", quietly = TRUE)) install.packages("SpatialKDE")
library(SpatialKDE)
```

```{r mask results 3, fig.width = 5.5, fig.height = 5, eval = FALSE}
# Note: this code is not evaluated when building the vignette as it requires the SpatialKDE package

# Spatial KDE requires an sf data.frame containing only POINTS that is in a projected coordinate system
# The simulated coordinates are not projected, but for the purpose of this example, we'll pretend they are projected under the Goode Homolosine projection
# This kind of arbitrary setting of crs should not be done for real data (see above example for how to properly project coordinates)
lotr_sf <- st_as_sf(lotr_coords, coords = c("x", "y")) %>% st_set_crs("+proj=goode")

# The grid layer must also be a RasterLayer for SpatialKDE
# Here, we use the kriged raster as our grid to get a KDE layer of the same spatial extent and resolution
grid <- raster(kgd)

# See the SpatialKDE package for more options and details about using KDE
kde_lyr <- kde(
  lotr_sf,
  kernel = "quartic",
  band_width = 15,
  decay = 1,
  grid = grid,
)

# Visualize KDE layer
ggplot_count(kde_lyr) +
  ggtitle("KDE")

# Mask with mask_gd
mgd <- mask_gd(kgd, kde_lyr, minval = 1)

ggplot_gd(mgd) +
  ggtitle("Kriged & sample density masked pi")
```

Another nice visualization option is to add a "background" to your plots
in the form of a raster or other sp or sf object (e.g., a country or
range boundary) which can help provide geographic context:

```{r range map background, fig.width = 5, fig.height = 5, warning = FALSE}
ggplot_gd(mgd, bkg = lotr_range) +
  ggtitle("Kriged & masked pi")
```

## Parallelization

To increase computational speed, users can perform any of the moving
window functions calculations with parallelization using future::plan().

Checkout the [furrr package
documentation](https://furrr.futureverse.org/) for more details on
setting up the back end for parallelization.

```{r parallelization, fig.width = 5, fig.height = 5, eval = FALSE, message = FALSE}
# Note: this code is not evaluated when building the vignette as it spawns multiple processes

# setup parallel session
future::plan("multisession", workers = 2)

wgd <- window_gd(lotr_vcf,
  lotr_coords,
  lotr_lyr,
  stat = "pi",
  wdim = 7,
  fact = 3,
  rarify_n = 2,
  rarify_nit = 5,
  rarify = TRUE
)

# end parallel session
future::plan("sequential")
```

*Note:* the `parallel` and `ncores` arguments have been deprecated as of
wingen 2.0.1.

## Other genetic diversity metrics

You can calculate multiple statistics at once by providing a vector of
statistic names to `stat`.

```{r, cache = TRUE, warning = FALSE, warning = FALSE, message = FALSE}
multistat_wgd <- window_gd(lotr_vcf,
  lotr_coords,
  lotr_lyr,
  stat = c("pi", "Ho"),
  wdim = 7,
  fact = 3,
  rarify = FALSE,
  L = 100
)
```

```{r, fig.width = 5, fig.height = 5, warning = FALSE}
ggplot_gd(multistat_wgd, bkg = lotr_range)
```

`"hwe"` and `"basic_stats"` take longer to calculate, so we do not evaluate these statistics here. We also change the parameters and reduce the dataset in the example to make it faster for users who wish to run it:

```{r, warning = FALSE, message = FALSE, eval = FALSE}
hwe_wgd <- window_gd(lotr_vcf[1:10, ],
  lotr_coords,
  lotr_lyr,
  stat = "hwe",
  wdim = 3,
  fact = 5,
  rarify = FALSE,
  L = 100,
  sig = 0.10
)

bs_wgd <- window_gd(lotr_vcf[1:10, ],
  lotr_coords,
  lotr_lyr,
  stat = "basic_stats",
  wdim = 3,
  fact = 5,
  rarify = FALSE,
  L = 100
)

ggplot_gd(hwe_wgd, bkg = lotr_range)

ggplot_gd(bs_wgd, bkg = lotr_range)
```

# General moving window

We provide a `window_general` function that can be used to make moving
window maps for other types of data inputs and functions. Unlike
`window_gd`, `window_general` does not require a `vcfR` object or a path
to a vcf file as input.

The required input (`x`) depends on the statistic (`stat`) being
calculated.

For the standard genetic diversity statistics:

-   If `stat = "pi"` or `"biallelic_richness"`, `x` must be a dosage
    matrix with values of 0, 1, or 2 (*Note:* rows must be individuals).

-   If `stat = "Ho"`, `x` must be a heterozygosity matrix where values
    of 0 = homozygosity and values of 1 = heterozygosity (*Note:* rows
    must be individuals).

-   If `stat = "allelic_richness"`, `x` must be a `genind` type object.

-   If `stat = "basic_stats"`, `x` must be a `hierfstat` type object.

For other statistics:

-   If `x` is a vector, `stat` can be any function that can be applied
    to a vector (e.g., `stat = mean, var, sum, etc.`).

-   If `x` is a matrix or data frame (*Note:* rows must be individuals),
    `stat` can be any function that takes a matrix or data frame and
    outputs a single numeric value (e.g., a function that produces a
    custom diversity index). *(Note: this functionality has not have
    been tested extensively and may produce errors, so use with
    caution)*.

As an example, let's create a moving window map of our raster layer
values (e.g., carrying capacity and conductance, in this case) at the
sample coordinates:

```{r, fig.width = 5, fig.height = 5, cache = TRUE, warning = FALSE, message = FALSE}
# First, we extract the raster values at those coordinates
vals <- extract(lotr_lyr, lotr_coords)

# Next, we run the window_general function with the env vector and set the `stat` to mean
# Note: we can also provide additional arguments to functions, such as na.rm = TRUE
we <- window_general(vals,
  coords = lotr_coords,
  lyr = lotr_lyr,
  stat = mean,
  wdim = 7,
  fact = 3,
  rarify_n = 2,
  rarify_nit = 5,
  rarify = TRUE,
  na.rm = TRUE
)

ggplot_gd(we) +
  ggtitle("Mean raster value")
```

# Other genetic data input file types in wingen

Although the default input data type for wingen is a VCF file, which is
a standard file type to encode genomic data, wingen can accept a genind
object as input to calculate allelic richness using `window_general()`.
Given that some wingen functionality can be feasible using a genind
object, users could easily convert various other genetic data file types
to then run wingen.

For example, genepop (.gen) is a typical file format for encoding
microsatellite data, in which each allele within a locus is coded using
a two- or three-digit system (i.e., in a diploid organism, in a
two-digit system, each locus would be assigned four digits specifying
each of two alleles. In a three-digit system in the same organism, each
locus would be assigned six digits encoding those same two alleles).
Users can convert a genepop file into a genind object using the
`read.genepop()` function in the adegenet package, which automatically
reads in a .gen file as a genind object. In many cases, microsatellite
data may not be biallelic, and may contain more than one observed allele
at a given locus; in such cases, running `window_general()` with a
genind object is still feasible.

An example of how one could use a genind object as input into
`window_general()` is as follows:

```{r, fig.width = 5.5, fig.height = 5, cache = TRUE, warning = FALSE, message = FALSE, eval = FALSE}
# We use the vcfR package to convert vcf to genind for our example
library(vcfR)

# Convert existing vcf example file into genind object:
genind <- vcfR2genind(lotr_vcf)

# Run window_general with no rarefaction
we_gi <- window_general(genind,
  coords = lotr_coords,
  lyr = lotr_lyr,
  stat = "allelic_richness",
  wdim = 7,
  fact = 3,
  na.rm = TRUE
)
```
