---
title: "Empirical Example Figures"
output: github_document
---
```{r, message = FALSE, results = FALSE, warning = FALSE, include = FALSE}
library(wingen)
library(vcfR)
library(rgdal)
library(raster)
library(viridis)
library(here)
library(tigris)
library(sf)
library(purrr)
library(terra)

wdir <- here("paperex", "empex")
source(here(wdir, "empex_functions.R"))

# create plot directory
dir.create(here(wdir, "plots"))

# temporary:
devtools::load_all()
```

# Empirical Dataset

Data used in this example is from Bouzid et al. (2022): https://doi.org/10.5061/dryad.n5tb2rbv2

**Bouzid, N. M., Archie, J. W., Anderson, R. A., Grummer, J. A., & Leaché, A. D. (2022). Evidence for ephemeral ring species formation during the diversification history of western fence lizards ( Sceloporus occidentalis ). Molecular Ecology, 31(2), 620–631. https://doi.org/10.1111/mec.15836**

```{r, fig.width = 5, fig.height = 5}

# Genetic data
vcf <- read.vcfR(here(wdir, "data", "populations_r20.haplotypes.filtered_m70_randomSNP.vcf"))

# Coordinates
coords <- read.table(here(wdir, "data", "Scelop.coord"))
coords_longlat <- st_as_sf(coords, coords = c("V1", "V2"), crs = "+proj=longlat") 
coords_proj <- st_transform(coords_longlat, crs = 3310)

```

Additionally, state data is used from TIGRIS:


```{r, message = FALSE, results = FALSE}
# download states from tigris
states <- states(cb = TRUE)

# reproject into wgs84 to match coordinates
states_longlat <- st_transform(states, crs = "+proj=longlat")
states_proj <- st_transform(states, crs = 3310)

# subset out CONUS
conus <- states_longlat[-which(states_longlat$NAME %in% c("Alaska", "Hawaii", "Puerto Rico", "American Samoa", "Guam", "Commonwealth of the Northern Mariana Islands", "United States Virgin Islands")), "STUSPS"]
# convert to SPDF for plotting
conus <- as_Spatial(conus)

# subset out Northern US
NUS_proj <- states_proj[which(states_proj$NAME %in% c("California", "Oregon", "Washington", "Nevada", "Idaho")), "STUSPS"]
NUS_longlat <- states_longlat[which(states_longlat$NAME %in% c("California", "Oregon", "Washington", "Nevada", "Idaho")), "STUSPS"]
```


### **Figure 4:** Geographic context plots

```{r, fig.width = 8, fig.height = 5}
pdf(here(wdir, "plots", "figure_4.1.pdf"))
par(mar = rep(0,4))
plot(conus, col = "lightgray", border = "lightgray", main = "")
plot(NUS_longlat, col = mako(1, begin = 0.7), border = "white", add = TRUE, main = "")
dev.off()
```

```{r, fig.width = 3, fig.height = 5}
pdf(here(wdir, "plots", "figure_4.2.pdf"))
par(mar = rep(0,4))
plot(NUS_longlat, col = mako(1, begin = 0.7), border = "white", lwd = 3, main = "")
dev.off()
```

## Run wingen analysis

First, different parameter combinations are evaluated. Here we vary the window size (wdim), the raster resolution (res), and the rarefaction size (rarify_n) to get the plots from Figure S4

### **Figure S5** Comparison of parameters

```{r, fig.width = 7, fig.height = 8, cache = TRUE, result = FALSE, message = FALSE}
params <- df_to_ls(expand.grid(res = c(30000, 40000, 50000), wdim = c(3, 5), rarify_n = c(2, 3, 4)))

stk <- map(params, test_params_empex, vcf, coords_proj)

stk_longlat <- map(stk, ~terra::project(.x, "+proj=longlat"))

par(mfrow = c(2, 3), mar = rep(1, 4))
walk(stk_longlat, test_empex_plot, bkg = NUS_longlat, zlim = c(0.02, 0.11))
```

Based on the results above we chose a final set of parameters:

```{r}
# set parameters 
wdim = 5
fact = 0
res = 30000

# create final layer
lyr <- coords_to_raster(coords_proj, res = res, buffer = 11)
```

And ran the moving window function again, this time kriging and masking the resulting rasters:

```{r, fig.width = 5, fig.height = 5, cache = TRUE, message = FALSE, results = FALSE, warning = FALSE}
# Run moving window
set.seed(22)
st <- Sys.time()
pg <- window_gd(vcf, coords_proj, lyr, stat = "pi", wdim = wdim, fact = fact, rarify = TRUE, rarify_n = 2, rarify_nit = 5)
pg_longlat <- terra::project(pg, "+proj=longlat")
Sys.time() - st

plot_gd(pg)
set.seed(22)
st <- Sys.time()
ag <- window_gd(vcf, coords_proj, lyr, stat = "biallelic_richness", wdim = wdim, fact = fact, rarify = TRUE, rarify_n = 2, rarify_nit = 5)
ag_longlat <- terra::project(ag, "+proj=longlat")
Sys.time() - st

set.seed(22)
st <- Sys.time()
hg <- window_gd(vcf, coords_proj, lyr, stat = "Ho", wdim = wdim, fact = fact, rarify = TRUE, rarify_n = 2, rarify_nit = 5)
hg_longlat <- terra::project(hg, "+proj=longlat")
Sys.time() - st


# krige and mask layers
mpg <- empex_krig_mask(pg, lyr, NUS_longlat)
mag <- empex_krig_mask(ag, lyr, NUS_longlat)
mhg <- empex_krig_mask(hg, lyr, NUS_longlat)
```

### **Figure 4:** Comparison of different measures: 

```{r, fig.width = 23, fig.height = 10}
colpal <- colorRampPalette(c("white", "#EFEDF5", "#DADAEB","#BCBDDC", "#9E9AC8", "#807DBA", "#6A51A3", "#54278F"))
pdf(here(wdir, "plots", "figure_4.3.pdf"), width = 24, height = 10)
par(mfrow = c(1,3), mar = rep(1,4), oma = rep(2.5,4))
raster_plot_gd(mpg, legend = FALSE, col = colpal(20))
plot(NUS_longlat, add = TRUE, col = NA, border = "#808080", lwd = 3)
points(coords, pch = 16, col = "black", cex = 2.1)
raster_plot_gd(mpg, smallplot = c(0.0, .02, .2,.75), legend.width = 2, legend.shrink = 0.1, axis.args = list(cex.axis = 3), legend.only = TRUE, add = TRUE, col = colpal(20))

raster_plot_gd(mag, legend = FALSE, col = colpal(20))
plot(NUS_longlat, add = TRUE, col = NA, border = "#808080", lwd = 3)
points(coords, pch = 16, col = "black", cex = 2.1)
raster_plot_gd(mag, smallplot = c(0.0, .02, .2,.75), legend.width = 2, legend.shrink = 0.1, axis.args = list(cex.axis = 3), legend.only = TRUE, add = TRUE, col = colpal(20))

raster_plot_gd(mhg, legend = FALSE, col = colpal(20))
plot(NUS_longlat, add = TRUE, col = NA, border = "#808080", lwd = 3)
points(coords, pch = 16, col = "black", cex = 2.1)
raster_plot_gd(mhg, smallplot = c(0.0, .02, .2,.75), legend.width = 2,legend.shrink = 0.1, axis.args = list(cex.axis = 3), legend.only = TRUE, add = TRUE, col = colpal(20))
dev.off()

pdf(here(wdir, "plots", "figure_4.4.pdf"))
par(mfrow = c(1,3), mar = rep(1.8,4), oma = rep(2.5,4))
raster_plot_gd(mpg, legend.width = 2, legend.shrink = 0.5, axis.args = list(cex.axis = 2), legend.only = TRUE)

raster_plot_gd(mag, legend.width = 2, legend.shrink = 0.5, axis.args = list(cex.axis = 2), legend.only = TRUE)

raster_plot_gd(mhg, legend.width = 2,legend.shrink = 0.5, axis.args = list(cex.axis = 2), legend.only = TRUE)
dev.off()
```


### **Figure S6:**  moving window calculations with and without rarefaction 

```{r, fig.width = 5, fig.height = 5, cache = TRUE}
set.seed(22)
hgn <- window_gd(vcf, coords_proj, lyr, stat = "Ho", wdim = wdim, fact = fact, rarify = FALSE, min_n = 2)
hgn_longlat <- terra::project(hgn, "+proj=longlat")

set.seed(22)
pgn <- window_gd(vcf, coords_proj, lyr, stat = "pi", wdim = wdim, fact = fact, rarify = FALSE, min_n = 2, L = nrow(vcf))
pgn_longlat <- terra::project(pgn, "+proj=longlat")

set.seed(22)
agn <- window_gd(vcf, coords_proj, lyr, stat = "biallelic_richness", wdim = wdim, fact = fact, rarify = FALSE, min_n = 2, rarify_alleles = TRUE)
agn_longlat <- terra::project(agn, "+proj=longlat")
```


```{r, fig.width = 15, fig.height = 12}

par(mfrow = c(2, 3), mar = rep(1,4), oma = rep(3,4))

raster_plot_gd(pg_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(ag_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hg_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))

raster_plot_gd(pgn_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(agn_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hgn_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))

par(mfrow = c(2, 3), mar = rep(0,4), oma = c(0, 0, 0, 3))

raster_plot_gd(pg_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(pg, pgn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(ag_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(ag, agn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hg_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(hg, hgn), legend.width = 2, axis.args = list(cex.axis = 2))

raster_plot_gd(pgn_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(pg, pgn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(agn_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(ag, agn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hgn_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(hg, hgn), legend.width = 2, axis.args = list(cex.axis = 2))
```

```{r, fig.width = 4, fig.height = 5}
par(mar = rep(0,4), oma = rep(0,4))
raster_plot_count(ag[[2]], legend.width = 1, col = mako(100))
```

These plots were then edited in Adobe Illustrator to create the final figures
