---
title: "Empirical Example"
output: github_document
---
```{r, message = FALSE, results = FALSE, warning = FALSE, include = FALSE}
library(wingen)
library(vcfR)
library(rgdal)
library(raster)
library(viridis)
library(here)
library(tigris)
library(sf)
library(purrr)

wdir <- here("paperex", "empex")
source(here(wdir, "empex_functions.R"))

# temporary:
devtools::load_all()
```

# Empirical Dataset

Data used in this example is from [Bouzid et al. (2022)](https://doi.org/10.5061/dryad.n5tb2rbv2)

**Bouzid, N. M., Archie, J. W., Anderson, R. A., Grummer, J. A., & Leaché, A. D. (2022). Evidence for ephemeral ring species formation during the diversification history of western fence lizards (Sceloporus occidentalis). Molecular Ecology, 31(2), 620–631. https://doi.org/10.1111/mec.15836**


The vcf used in this example was created from the ped and map file from Bouzid et al. 2022 using Plink (this is not run as part of the notebook, but we provide the code below).

```bash
./plink --file populations_r20.haplotypes.filtered_m70_randomSNP --recode vcf-fid --out populations_r20.haplotypes.filtered_m70_randomSNP
```

```{r, fig.width = 5, fig.height = 5}
# Genetic data
vcf <- read.vcfR(here(wdir, "data", "populations_r20.haplotypes.filtered_m70_randomSNP.vcf"))

# Coordinates
coords <- read.table(here(wdir, "data", "Scelop.coord"))
coords_longlat <- st_as_sf(coords, coords = c("V1", "V2"), crs = "+proj=longlat") 
coords_proj <- st_transform(coords_longlat, crs = 3085)
plot(coords_proj)
```

Additionally, state data is used from TIGRIS:

```{r, message = FALSE, results = FALSE}
# download states from tigris
states <- states(cb = TRUE)

# reproject into wgs84 to match coordinates
states_longlat <- st_transform(states, crs = "+proj=longlat")
states_proj <- st_transform(states, crs = 3085)

# subset out CONUS
conus <- states_longlat[-which(states_longlat$NAME %in% c("Alaska", "Hawaii", "Puerto Rico", "American Samoa", "Guam", "Commonwealth of the Northern Mariana Islands", "United States Virgin Islands")), "STUSPS"]
# convert to SPDF for plotting
conus <- as_Spatial(conus)

# subset out Northern US
NUS_proj <- states_proj[which(states_proj$NAME %in% c("California", "Oregon", "Washington", "Nevada", "Idaho")), "STUSPS"]
NUS_longlat <- states_longlat[which(states_longlat$NAME %in% c("California", "Oregon", "Washington", "Nevada", "Idaho")), "STUSPS"]

plot(NUS_proj)
```

### **Figure 4:** Geographic context plots

```{r, fig.width = 8, fig.height = 5}
par(mar = rep(0,4))
plot(conus, col = "lightgray", border = "lightgray", main = "")
plot(NUS_longlat, col = mako(1, begin = 0.7), border = "white", add = TRUE, main = "")
```

```{r, fig.width = 3, fig.height = 5}
par(mar = rep(0,4))
plot(NUS_longlat, col = mako(1, begin = 0.7), border = "white", lwd = 2, main = "")
```

## Run wingen analysis

First, different parameter combinations are evaluated. Here, we vary the window size (`wdim`), the raster resolutions (`res`), and the rarefaction size (`rarify_n`).

### **Figure S5** Comparison of parameters

```{r, fig.width = 7, fig.height = 8, cache = TRUE, result = FALSE, message = FALSE}
params <- df_to_ls(expand.grid(res = c(30000, 40000, 50000), wdim = c(3, 5), rarify_n = c(2, 3, 4)))

stk <- map(params, test_params_empex, vcf, coords_proj)

stk_longlat <- map(stk, ~terra::project(.x, "+proj=longlat"))

par(mfrow = c(2, 3), mar = rep(1, 4))
walk(stk_longlat, test_empex_plot, bkg = NUS_longlat, zlim = c(0.02, 0.11))
```

Based on the results above, we chose a final set of parameters:

```{r}
# set parameters 
wdim = 5
fact = 0
res = 30000

# Create final layer 
lyr <- coords_to_raster(coords_proj, res = res, buffer = 10)

plot_gd(lyr, breaks = 100)
```

And ran the moving window function again, this time kriging and masking the resulting rasters:

```{r, fig.width = 5, fig.height = 5, cache = TRUE, message = FALSE, results = FALSE, warning = FALSE}
# Run moving window
set.seed(22)
st <- Sys.time()
pg <- window_gd(vcf, coords_proj, lyr, stat = "pi", wdim = wdim, fact = fact, rarify = TRUE, rarify_n = 2, rarify_nit = 5)
pg_longlat <- terra::project(pg, "+proj=longlat")
Sys.time() - st

plot_gd(pg)
set.seed(22)
st <- Sys.time()
ag <- window_gd(vcf, coords_proj, lyr, stat = "biallelic_richness", wdim = wdim, fact = fact, rarify = TRUE, rarify_n = 2, rarify_nit = 5)
ag_longlat <- terra::project(ag, "+proj=longlat")
Sys.time() - st

set.seed(22)
st <- Sys.time()
hg <- window_gd(vcf, coords_proj, lyr, stat = "Ho", wdim = wdim, fact = fact, rarify = TRUE, rarify_n = 2, rarify_nit = 5)
hg_longlat <- terra::project(hg, "+proj=longlat")
Sys.time() - st

# krige and mask layers
mpg <- empex_krig_mask(pg, lyr, NUS_longlat)
mag <- empex_krig_mask(ag, lyr, NUS_longlat)
mhg <- empex_krig_mask(hg, lyr, NUS_longlat)
```

### **Figure 4:** Comparison of different measures: 

```{r, fig.width = 12, fig.height = 6}
colpal <- colorRampPalette(c("white", "#EFEDF5", "#DADAEB","#BCBDDC", "#9E9AC8", "#807DBA", "#6A51A3", "#54278F"))
par(mfrow = c(1,3), mar = rep(2,4), oma = rep(2.5,4))
raster_plot_gd(mpg, legend.width = 2, axis.args = list(cex.axis = 1.5), col = colpal(20))
plot(NUS_longlat, add = TRUE, col = NA, border = "#808080", lwd = 2)
plot(coords_longlat, pch = 16, col = "black", cex = 1.5, add = TRUE)

raster_plot_gd(mag, legend.width = 2, axis.args = list(cex.axis = 1.5), col = colpal(20))
plot(NUS_longlat, add = TRUE, col = NA, border = "#808080", lwd = 2)
plot(coords_longlat, pch = 16, col = "black", cex = 1.5, add = TRUE)

raster_plot_gd(mhg, legend.width = 2, axis.args = list(cex.axis = 1.5), col = colpal(20))
plot(NUS_longlat, add = TRUE, col = NA, border = "#808080", lwd = 2)
plot(coords_longlat, pch = 16, col = "black", cex = 1.5, add = TRUE)
```

### **Figure S6:**  moving window calculations with and without rarefaction 

```{r, fig.width = 5, fig.height = 5, cache = TRUE}
set.seed(22)
pgn <- window_gd(vcf, coords_proj, lyr, stat = "pi", wdim = wdim, fact = fact, rarify = FALSE, min_n = 2, L = nrow(vcf))
pgn_longlat <- terra::project(pgn, "+proj=longlat")

set.seed(22)
agn <- window_gd(vcf, coords_proj, lyr, stat = "biallelic_richness", wdim = wdim, fact = fact, rarify = FALSE, min_n = 2, rarify_alleles = TRUE)
agn_longlat <- terra::project(agn, "+proj=longlat")

set.seed(22)
hgn <- window_gd(vcf, coords_proj, lyr, stat = "Ho", wdim = wdim, fact = fact, rarify = FALSE, min_n = 2)
hgn_longlat <- terra::project(hgn, "+proj=longlat")

```


```{r, fig.width = 15, fig.height = 12}

par(mfrow = c(2, 3), mar = rep(1,4), oma = rep(3,4))

raster_plot_gd(pg_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(ag_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hg_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))

raster_plot_gd(pgn_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(agn_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hgn_longlat, NUS_longlat, breaks = 100, legend.width = 2, axis.args = list(cex.axis = 2))

par(mfrow = c(2, 3), mar = rep(1,4), oma = rep(3,4))

raster_plot_gd(pg_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(pg, pgn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(ag_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(ag, agn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hg_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(hg, hgn), legend.width = 2, axis.args = list(cex.axis = 2))

raster_plot_gd(pgn_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(pg, pgn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(agn_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(ag, agn), legend.width = 2, axis.args = list(cex.axis = 2))
raster_plot_gd(hgn_longlat, NUS_longlat, breaks = 100, zlim = get_minmax(hg, hgn), legend.width = 2, axis.args = list(cex.axis = 2))
```

```{r, fig.width = 4, fig.height = 5}
par(mar = rep(0,4))
raster_plot_gd(ag_longlat[[2]], col = mako(100))
```

test
